### 2011-01-28 trunk/papers/webist2011/paper_ssec_discovery.tex
power shortening and improving
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1208 d8b0c59b-6544-0410-94d9-83a5658a058e

\subsection{Discovery mechanisms} % (fold)
\label{ssec:discovery}

The platform presented in this paper is intended to support web service discovery and recommendation capabilities.
As explained in detail in Section~\ref{ssec:ontology}, there are different types of building blocks which can be modelled
using the platform, leading to a necessity for different kinds of discovery or recommendation mechanisms, depending
on the level of composition to deal with: screen-flow or screen. In the screen-flow composition, the smallest
functional unit is the screen, meaning that a screen-flow is composed by interconnecting a number of screens between 
each other. This composition can be applied to any sort of building block since they all share the
same structure: a set of pre- and post-conditions, which make possible the communication with other building blocks.
That said, technically the algorithms presented in this section could be used for the discovery and composition of
any type of building block stored in the platform.

At the screen-flow design level, the composition is made using a set of screens, and the pre- and post-conditions 
of the screen-flow itself used as entry and exit points of the screen-flow. A screen is only reachable when its 
pre-conditions are satisfied as well, meaning that the post-conditions of another screen, or the pre-conditions of 
the screen-flow are compatible with them.

The main goal of the discovery process is to help the user to make the screen-flow executable, in other words, to make
all the screens within a screen-flow reachable. The platform offers two mechanisms to find and recommend
screens, previously stored in the building block base (the catalogue), which will make the screen-flow executable: a
simple discovery based on pre- and postconditions, and a multi-step discovery or planning algorithm. Like any other
search engine or recommender system, the results are ranked before being presented
to the user. The ranking algorithm is covered in Section~\ref{sssec:ranking}.

\subsubsection{Simple discovery based on pre- and post-conditions}
\label{sssec:simple_discovery}

At a particular state of a composition, ie. at screen-flow design, several of the building blocks used to compose it, 
screens in the case of a screen-flow, might be unreachable, therefore there must be pre-conditions not satisfied. The
platform will assist the process by recommending building blocks which will satisfy these pre-conditions. In a nutshell,
it collects the pre-conditions of all the unreachable building blocks, used as a graph pattern (see Section~\ref{ssec:ontology})
in order to be matched against the post-conditions of any building block. In the following scenario, there are two 
screens: \emph{s1} and \emph{s2}. \emph{s1} has as a pre-condition: ``there exists a search criteria'', and as a 
post-condition: ``there exists a item''. \emph{s2} has just a pre-condition stating: ``there exists a search criteria''.
Here you can see the same example but formally defined:

>>>>

\subsection{Discovery Mechanisms} % (fold)
\label{ssec:discovery}

% Due to the diversity of building blocks available, there is a necessity for different kinds of discovery or recommendation mechanisms, depending on the level of composition: screen-flow or screen.

The main goal of the discovery process is to aid the user in finding suitable building blocks to complement the ones they are already using. E.g., on the screen-flow level, this would mean to suggest additional screens to make existing screens within a screen-flow reachable, and the screen-flow therefore executable.
The platform offers two mechanisms to find and recommend screens (or other building blocks) stored in the catalogue: a simple discovery based on pre- and postconditions, and a multi-step discovery or planning algorithm. 
Before being presented to the user, the results are being ranked, as discussed in Sect.~\ref{sssec:ranking}.

\subsubsection{Simple Discovery Based on Pre- and Post-Conditions}
\label{sssec:simple_discovery}

In this simple approach, the platform will assist the process by recommending all building blocks which will satisfy currently unfulfilled pre-conditions. 
The pre-conditions of all the unreachable building blocks are collected as a graph pattern, which is then matched against the post-conditions of all available building blocks. 
In the following scenario, there are two 
screens: \emph{s1} and \emph{s2}. \emph{s1} has as a pre-condition: \emph{``there exists a search criteria''}, and as a 
post-condition: \emph{``there exists a item''}. \emph{s2} has just a pre-condition stating: \emph{``there exists a search criteria''}.

### 2011-01-28 trunk/papers/webist2011/paper_ssec_ontology.tex
power shortening and improving
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1208 d8b0c59b-6544-0410-94d9-83a5658a058e

\subsection{Conceptual Model}
\label{ssec:ontology}
The conceptual model used to define the web service wrappers within this application has been influenced by WSDL, the main web service definition language, and semantic approaches such as WSMO and OWL-S. It belongs to a more complex conceptual model for FAST, which can be grouped into three levels: the gadget and screenflow level at the top, the level of individual screens in the middle, and the level of web services at the bottom. 
In short, a screenflow is an executable (in a Web browser) set of screens, while a screen is a set of building blocks with a specific purpose. These building blocks --- \emph{forms} for the UI, \emph{operators} for data manipulation and \emph{back-end service} --- all share the same structure. Every building block has a set of actions (operations in WSDL), each of which needs a set of pre-conditions (inputs) to be fulfilled in order to be executed, and provides a set of post-conditions (outputs) to other building blocks. These pre- and post-conditions are defined as individual RDF triples. This means that conditions can be modelled as graph patterns using SPARQL notation \cite{sparql2008spec}, where e.g. the post-condition of a login service such as ``there exists a user'' will be transformed into a simple pattern like ``a variable \texttt{?user} is of type \texttt{sioc:User}'', or ``\texttt{?user a sioc:User}''. Using this mechanism, extended by RDFS entailment rules, services and other building blocks can be matched automatically. The publishing and discovery platform (catalogue) can employ this functionality to support the discovery of web services based on the current user needs (expressed in the same way as pre-conditions). For further details on the conceptual model, see its definition in \cite{moeller2010fast_ontology}.

>>>>

\subsection{Conceptual Model}
\label{ssec:ontology}
The conceptual model used to define the web service wrappers within this application has been influenced by both WSDL and semantic approaches such as WSMO and OWL-S. 
It is part of a more complex conceptual model for FAST~\cite{moeller2010fast_ontology}, which can be grouped into three levels: the gadget and screenflow level at the top, the level of individual screens in the middle, and the level of web services at the bottom. 
All these building blocks and their sub-parts
% --- \emph{forms} for the UI, \emph{operators} for data manipulation and \emph{back-end services} --- 
share the same structure: a set of actions (like operations in WSDL), each of which needs a set of pre-conditions (inputs) to be fulfilled in order to be executed, and provides a set of post-conditions (outputs) to other building blocks. These pre- and post-conditions are defined as RDF graph patterns. E.g., the post-condition of a login service such as ``there exists a user'' will be expressed as a simple pattern such as ``\texttt{?user a sioc:User}''\footnote{We use Trig (\url{http://www4.wiwiss.fu-berlin.de/bizer/TriG/}) and SPARQL (\url{http://www.w3.org/TR/rdf-sparql-query/}) notation for RDF graphs throughout this paper.}. Using this mechanism, extended by RDFS entailment rules, services and other building blocks can be matched automatically. The publishing and discovery platform can employ this functionality to support the internal discovery of web services based on the current user needs (expressed in the same way as pre-conditions).
